Leadership across cultures and contexts: an exploratory analysis

### 2020-05-07 leadership_across_cultures_contexts.Rmd

```{r elastichighstatus, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *High Status*. A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *High status* in each text record. B: Non-zero coefficients of all quality and function variables that best predicted evidence for *High status*. Both models used the lasso penality ($\\alpha=1$), with $\\lambda_{1se}$ chosen by cross-validation (1 SE from $\\lambda_{min}$). Red values are positive predictors; purple values are negative predictors. Coefficents = 0 not displayed. X-axes are odds ratios on a log 10 scale."}

@@

>>>>

```{r highstatusplot, fig.width=10, fig.cap="Coefficients of logistic elasticnet regression models [@friedman_regularization_2010] of evidence for *High Status*. A: Coefficients indicate the words whose frequencies in each text record best predicted evidence for *High status* in each text record. B: Non-zero coefficients of all quality and function variables that best predicted evidence for *High status*. Both models used the lasso penality ($\\alpha=1$), with $\\lambda_{1se}$ chosen by cross-validation (1 SE from $\\lambda_{min}$). Red values are positive predictors; purple values are negative predictors. Variables with coefficents = 0 not displayed. X-axes are odds ratios on a log 10 scale."}

@@

We found strong evidence that several dimensions of leadership were universal across cultures, and across group contexts within cultures. Leaders function to resolve conflict, organization cooperation, and provide counsel and direction in over 70% of cultures, and are seen as high status, knowledgeable and intelligence, and experienced and accomplished in about 80% or more of cultures. See Figure \@ref(fig:varsupport). Leaders benefitted materially, reproductively, and socially in 50% or more of cultures. See Figure \@ref(fig:varsupportcb). Common leader dimensions that did not vary by continental region, subsistence strategy, group context, or leader sex included qualities, such as being knowledgeable and intelligence, wealthly, strategically nepotistic, and generous; and functions, such as resolving conflict and representing the group. These dimensions are therefore candidates for universal dimensions of human leadership. By "universal", we do not mean that all leaders have all these traits. Instead, we mean that there is a fairly high level of evidence for these traits across cultures that does not vary by continental reagion, subsistence strategy, group context, or leader sex. See Table \@ref(tab:leaderuniversals).

Evidence for many leadership dimensions did vary across cultures and contexts, however, including functions such as organizing cooperation, providing counsel and direction and prosocial investment; and qualities such as high status and aggressiveness. See Figure \@ref(fig:heatmapdrop1).

Most discussions of leadership in the ethnographic record pertained to male leaders of community-level political groups, with leadership in residential subgroups and kin groups relatively more important in hunter-gatherer and pastoralist societies, and supracommunity political groups more important in horticultural and agricultural societies. Leadership in all social contexts, including economic, military, and religious, was present across all subsistence strategies, however. See Figure \@ref(fig:grouptype). The relative unimportance of supracommunity leadership in hunter-gatherers and pastorlists, and the relative importance of leadership in residential subgroups, highlights a major difference with WEIRD societies that more closely resemble the agricultural societies in our sample that show the opposite pattern.
